Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2017-05-02 manuscript.tex
latest

In the previous section we only considered how $b$, $c$ and $d$ should respond in Grime's environmental extremes. Here we further explore the low frequency behavior of Eq. \eqref{eq:master} to determine which types can coexist in a constant environment, and then consider the full time-dependent behaviour of Eq. \eqref{eq:master} in a cyclical environment.

Stable coexistence is possible in our extended lottery model. Suppose that $j$ is better at securing territories ($c_j>c_i$) but that $i$ is better at producing propagules ($b_i>b_j$), with $d_i=d_j$.

 This is a version of the classic competition-colonization trade-off \citep{tilman_94,levins_71}: the competitor ($c$-specialist) leaves many territories unoccupied (low $L$) due to its poor colonization ability (low $b$), which the colonizer ($b$-specialist) can then exploit. A similar argument applies for coexistence between high-$c$ and low-$d$ specialists; a ``competition-longevity'' trade-off \citep{tilman_94}. These forms of co-existence require density dependence, and are not present in the classic lottery model. Mutual invasion is not possible between $b$- and $d$-specialists.

@@

The classic lottery model (Eq. 1) gives dramatically different co-existence predictions because expansion to carrying capacity occurs immediately at the start of the summer (Fig. \ref{fig:fluctuatingselection}d-f). As a result, coexistence requires that the winter survivor's $b$ must be about $5$ times smaller than predicted when we properly account for the growth in the abundance of each type using Eq. \eqref{eq:master} (keeping the other parameters the same). Previous models of the promotion of genetic variation via the storage effect  \citep{ellner_1994} similarly assume that the total number of offspring per iteration is constant, and would produce a similar error.

@@

\begin{figure}
\centering
\includegraphics[scale=0.7]{fluctuatingselection.pdf}
\caption{\label{fig:fluctuatingselection} Stable coexistence between $b$ and $d$ specialists in a fluctuating environment. (a) Birth and death rates seasonally alternate being nonzero (white for winter, green for summer). The $b$ specialist (black) has higher $b$ and $d$ ($b=0.5$, $d=0.2$) than the $d$ specialist ($b=0.217$, $d=0.1$) (blue). (b) Both types grow during the positive $b$ phase, and decline during the positive $d$ phase, but the $d$ specialist does so at a lower rate. Total height (blue+black) is population density $N/T$. (c) Summer favors the $b$ specialist, winter the $d$ specialist, and they stably coexist. (d-f) Same as (a-c) for the classic lottery model, where coexistence now requires $b\approx 0.0421$ for the $d$ specialist. For illustration, the propagule abundances are assumed to have the form $m_i=b_i(1-N/T)n_i$, reflecting non-directed dispersal.} 
\end{figure}

@@

In the introduction we mentioned the recurring difficulties with confounding selection and demography in population genetic inference. While we have not directly attempted to perform inference with our model, and Eq. \eqref{eq:master} may not be appropriate for particular inference problems, it seems that something similar (and hopefully more analytically tractable) to our extension of the lottery model is unavoidable because, fundamentally, selective births and deaths affect both  abundances and frequencies, not one or the other in isolation. Moreover, some aspects of allele frequency change are intrinsically density-dependent. In the classic lottery model, which as we have seen is essentially the Wright-Fisher model with overlapping generations, $b_i$ and $c_i$ are  equivalent in the sense that the number of territorial victories only depends on the product $b_i c_i$ (see ``Model''). This is no longer the case in our extension, where $b$ and $c$ specialists can co-exist. This ``colonization-competition trade-off'' is well known in the co-existence literature \citep{tilman_94}. It and similar forms of ``spatial co-existence'' in stable environments have previously been modeled either with Levin's  qualitative representation of competition \citep{levins_71,tilman_94}, as opposed to the quantitative $c$ of lottery competition, or with a more sophisticated treatment of space (non-uniform dispersal; \citealt{shmida_84,bolker_99}). In cyclical environments, polymorphisms can be stabilized by the bounded density effect which is  completely lost if there is an exclusive focus on allele frequencies \citep{yi_2013}. We leave the details of how our model might be applied in inference, including the crucial issue of its genetic drift predictions, for future work.

It is interesting to compare the predictions of the extended lottery model with earlier approaches, such as the $r$/$K$ scheme, where $r=b-d$ is the maximal, low-density growth rate \citep{pianka_1972}. Confusingly, the term ``$K$-selection'' sometimes refers generally to selection at high density \citep{pianka_1972}, encompassing both selection for higher saturation density \citep{macarthur_1967} and competitive ability \citep{gill_1974}. Contrary to an $r$/$K$ dichotomy, empirical studies have shown that maximal growth rate and saturation density (measured by abundance) are positively correlated, both between species/strains \citep{luckinbill_1979,kuno_1991,hendriks_2005,fitzsimmons_2010}, and as a result of experimental evolution \citep{luckinbill_1978,luckinbill_1979}. From the perspective of our model, this correlation is not surprising since the saturation density, which is determined by a balance between births and deaths, increases with $b$.

@@

There is support for a negative relationship between competitive success at high density and maximal growth rate \citep{luckinbill_1979}, consistent with an $r$/$K$ dichotomy. This could be driven by a tradeoff between individual size and reproductive rate. To avoid confusion with other forms of ''$K$-selection'', selection for competitive ability has been called ``$\alpha$-selection'' after the competition coefficients in the Lotka-Volterra equation \citep{gill_1974,case_1974,joshi_2001}. However, competitive success as measured by $\alpha$ (i.e. the per-capita effect of one genotype on another genotype's growth rate) is only partly determined by individual competitive ability --- in the presence of age-structured competition and territoriality, it also includes the ability of each genotype to produce contestants i.e. $b$ in our model. Our $c$ is strictly competitive ability only --- as such, changes in $c$ do not directly affect population density (the total number of territories occupied per iteration is $\Delta_+ N=U(1-e^{-L})$, which does not depend directly on the $c_i$). The clean separation of a strictly-relative $c$ parameter is particularly useful from an evolutionary genetics perspective, essentially embedding a zero-sum relative fitness trait within a non-zero-sum fitness model. This could have interesting applications for modeling the impacts of intra-specific competition on species extinction, for example due to clonal interference \citep{gerrish_1998,desai_2007} between $c$-strategists on the one hand, and $b$- and $d$- strategists on the other.

@@

Our realization of Grime's triangle (Fig.~4) differs from approaches which identify primary strategies as trait combinations which can co-exist \citep{bolker_99}, referring instead to the direction of adaptive trait evolution under different regimes of stress and disturbance, which is closer in spirit to Grime's arguments \citep{grime_1974,grime_1977}. In addition, we have not assumed any kind of trade-offs or pleitropy between $b$, $c$ and $d$, only constraints imposed by the environment on the order of magnitude of $b$ and $d$. As an example of a trade-off, corals which rapidly out-shade neighbors have a tall, branched morphology which is vulnerable to disturbances, and so, all else being equal, ideal environment $c$-strategists will suffer higher mortality from disturbances \citep{darling_2012}. Fig.~\ref{fig:axes} gives the same conclusion without invoking trade-offs; mutations which reduce disturbance vulnerability are essentially neutral under ideal conditions, leading to no improvements in mortality from disturbances, whereas $c$ will tend to increase over time. Thus, while trade-offs may amplify specialization, and are sometimes invoked to explain primary strategy schemes \citep{macarthur_1967,winemiller_1992,aerts_1999}, they are not necessary for it.

One limitation of our model as a general-purpose model of density-dependent selection is the restriction of competition to interference competition between juveniles for durable resources (lottery recruitment to adulthood), analogous to the ubiquitous assumption of viability selection in population genetics \citep[p. 45]{ewens_2004}. In some respects this is the complement of resource competition models, which restrict their attention to exploitation competition, typically without age structure \citep{tilman_1982}. In the particular case that resources are spatially localized (e.g. due to restricted movement through soils), resource competition and territorial acquisition effectively coincide, and in principle resource competition could be represented by a competitive ability $c$ (or conversely, $c$ should be derivable from resource competition). The situation is more complicated if the resources are well-mixed, since, in general, resource levels then need to be explicitly tracked. It seems plausible that explicit resource tracking may not be necessary when the focus is on the evolution of similar genotypes rather than the stable co-existence of widely differing species \citep{ram_2016}. We are not aware of any attempts to delineate conditions under which explicit resource tracking is unnecessary even if it is assumed that community structure is ultimately determined by competition for consumable resources. More work is needed connecting resource competition models to the density-dependent selection literature, since most of the former has to date been focused on narrower issues of the role of competition at low resource availability \citep{aerts_1999,davis_1998,tilman_2007}.

@@

@@

\section*{Appendix C: Direction of evolution under different environmental extremes}

In this Appendix we consider mutant invasion of a resident type $i$ (such that $N=n_i$) which is in equilibrium.

In a population with a single genotype $i$ which is in equilibrium ($\Delta n_i = 0$), $R_i=0$, $\overline{c}=c_i$, $A_i=(1-(1+L)e^{-L})/L$, and Eq. \eqref{eq:master} gives
\begin{equation}
b_i(1-e^{-L})/L-d_i=0,\label{eq:equil}
\end{equation}
which implies $L\approx b_i/d_i$ when $b_i/d_i\gg 1$ and $L\ll 1$ when $b_i/d_i\approx 1$. (Eq. \eqref{eq:Da}{eq:equil} can also be seen directly from Eq. \eqref{eq:growthsumuncoupled}). Now suppose that a new genotype $j$, which is initially rare, appears in the population. Then $A_j/R_j\ll 0$, $l_j\approx 0$ and $\overline{c}\approx c_i$, and so, from Eq. \eqref{eq:master}, the mutant lineage's fitness is
\begin{equation}
\Delta n_j/n_j \approx b_j \left(e^{-L}+R_j\frac{c_j}{c_i}\right)-d_j \label{eq:invad}
\end{equation}
where $R_j\approx (1-e^{-L})/\left(\frac{c_j}{c_i}+\frac{L-1-e^{-L}}{1-(1+L)e^{-L}}\right)$.

To see that coexistence is possible between $b$ and $c$ specialists, we check if $j$ will invade an $i$-dominated population, but $i$ will also invade a $j$-dominated population (``mutual invasion''). Suppose that $b_i$ is so large that $L\gg 1$ when $i$ is dominant, and $b_j$ is so small that $L\ll 1$ when $j$ is dominant. Then, combining Eqs. \eqref{eq:equil} and \eqref{eq:invad}, we find that $i$ invades $j$ because $b_i>b_j$, while $j$ invades $i$ provided that
\begin{equation}
b_jc_jR_j-b_i c_i A_i>0. \label{eq:jinvadcoex}
\end{equation}
Thus, coexistence occurs if $c_j/c_i$ is large enough.

We now consider the different environments from section ``Primary strategies and Grime's triangle''. In our representation of Grime's stressful environment, $b\ll 1$, $d\approx b$ and $L\ll 1$, and so Eq. \eqref{eq:invad} becomes $\Delta n_j/n_j \approx b_j-d_j$. Mutations which improve $b$ or $d$ by the same fraction $\epsilon$, such that $b_j = (1+ \epsilon)b_i$ or $d_j = (1- \epsilon)d_i$,  yield identical fitness benefits.

In our representation of Grime's ideal environment, $d\ll 1$, $b_i/d_i \gg 1$ and $L\gg 1$, and so Eq. \eqref{eq:invad} becomes 
\begin{equation}
\Delta n_j/n_j \approx \frac{b_j c_j}{c_j + L-1}-d_j\approx \frac{b_j c_j}{L}-d_j, \label{eq:highLinvad}
\end{equation}
where the last approximation follows from the fact that $c_j$ will be of similar magnitude to $c_i=1$ (ignoring mutations with very large effect sizes; incidentally, Eq. \eqref{eq:lottery} yields the same expression by assuming $L\rightarrow \infty$ even though this is obviously unrealistic for a rare mutant). Similarly to Grime's stressful environment, improving $b$, $c$, or $d$ by a fraction $\epsilon$ yields identical fitness benefits, even though they may have very different magnitudes, because the ancestral resident type's density-regulated birth rate $b_i c_i/L$ exactly balances its mortality $d_i$ in equilibrium. Biases in the direction of evolution of $b$, $c$ or $d$ must again follow from biases in mutation availability or trade-offs/pleiotropy. The same argument applies for high-$L$ interpretation of the stress regime.

@@

>>>>

In the previous section we only considered how $b$, $c$ and $d$ should respond in Grime's environmental extremes, based on invasion fitness. Here we further explore the low frequency behavior of Eq. \eqref{eq:master} to determine which types can coexist in a constant environment, and then consider the full time-dependent behaviour of Eq. \eqref{eq:master} in a cyclical environment.

In a constant environment, stable coexistence is possible in our extended lottery model. A $b$ specialist $i$ and $c$-specialist $j$ ($b_i>b_j$, $c_j>c_i$) can co-exist because then propagule density $L$ is frequency dependent, and so is the importance of competitive ability (Appendix D). This is a version of the classic competition-colonization trade-off \citep{tilman_94,levins_71}: the competitor ($c$-specialist) leaves many territories unoccupied (low $L$) due to its poor colonization ability (low $b$), which the colonizer ($b$-specialist) can then exploit. A similar situation holds for coexistence between high-$c$ and low-$d$ specialists; a ``competition-longevity'' trade-off \citep{tilman_94}. These forms of co-existence require density dependence (being mediated by $L$), and are not present in the classic lottery model. Coexistence is not possible between $b$- and $d$-specialists (Appendix D).

@@

The classic lottery model (Eq. 1) gives dramatically different co-existence predictions because expansion to carrying capacity occurs immediately at the start of the summer (Fig. \ref{fig:fluctuatingselection}d-f). As a result, coexistence requires that the winter survivor's $b$ must be about $5$ times smaller than required when we properly account for the growth in the abundance of each type using Eq. \eqref{eq:master} (keeping the other parameters the same). Previous models of the promotion of genetic variation via the storage effect  \citep{ellner_1994} similarly assume that the total number of offspring per iteration is constant, and would produce a similar error.

@@

\begin{figure}
\centering
\includegraphics[scale=0.7]{fluctuatingselection.pdf}
\caption{\label{fig:fluctuatingselection} Stable coexistence between $b$ and $d$ specialists in a fluctuating environment requires a much greater $b$ advantage in the classic lottery model compared to our density-dependent extension of it when population density is seasonally cyclical. (a) Birth and death rates seasonally alternate being nonzero (white for winter, green for summer). The $b$ specialist (black) has higher $b$ and $d$ ($b=0.5$, $d=0.2$) than the $d$ specialist ($b=0.217$, $d=0.1$) (blue). (b) Both types grow during the positive $b$ phase, and decline during the positive $d$ phase, but the $d$ specialist does so at a lower rate. Total height (blue+black) is population density $N/T$. (c) Summer favors the $b$ specialist, winter the $d$ specialist, and they stably coexist. (d-f) Same as (a-c) for the classic lottery model; the types no longer coexist. (g-i) Same as (d-f) where now $b = 0.0421$ for the $d$ specialist and the types coexist. For illustration, the propagule abundances are assumed to have the form $m_i=b_i(1-N/T)n_i$, reflecting non-directed dispersal.} 
\end{figure}

@@

It is interesting to compare the predictions of the extended lottery model with earlier approaches, such as the $r$/$K$ scheme, where $r=b-d$ is the maximal, low-density growth rate \citep{pianka_1972}. Confusingly, the term ``$K$-selection'' sometimes refers generally to selection at high density \citep{pianka_1972}, encompassing both selection for higher saturation density \citep{macarthur_1967} and competitive ability \citep{gill_1974}. Contrary to predictions of an $r$/$K$ trade-off, empirical studies have shown that maximal growth rate and saturation density (measured by abundance) are positively correlated, both between species/strains \citep{luckinbill_1979,kuno_1991,hendriks_2005,fitzsimmons_2010}, and as a result of experimental evolution \citep{luckinbill_1978,luckinbill_1979}. From the perspective of our model, this positive correlation is not surprising since the saturation density, which is determined by a balance between births and deaths, increases with $b$.

@@

There is support for a negative relationship between competitive success at high density and maximal growth rate \citep{luckinbill_1979}, consistent with a tradeoff between $r$ and the competitive aspect of $K$. This could be driven by a tradeoff between individual size and reproductive rate. To avoid confusion with other forms of ''$K$-selection'', selection for competitive ability has been called ``$\alpha$-selection'' after the competition coefficients in the Lotka-Volterra equation \citep{gill_1974,case_1974,joshi_2001}. However, competitive success as measured by $\alpha$ (i.e. the per-capita effect of one genotype on another genotype's growth rate) is only partly determined by individual competitive ability --- in the presence of age-structured competition and territoriality, it also includes the ability of each genotype to produce contestants i.e. $b$ in our model. Our $c$ is strictly competitive ability only --- as such, changes in $c$ do not directly affect population density (the total number of territories occupied per iteration is $\Delta_+ N=U(1-e^{-L})$, which does not depend directly on the $c_i$). The clean separation of a strictly-relative $c$ parameter is particularly useful from an evolutionary genetics perspective, essentially embedding a zero-sum relative fitness trait within a non-zero-sum fitness model. This could have interesting applications for modeling the impacts of intra-specific competition on species extinction, for example due to clonal interference \citep{gerrish_1998,desai_2007} between $c$-strategists on the one hand, and $b$- and $d$- strategists on the other.

@@

Our realization of Grime's triangle (Fig.~4) differs from approaches which identify primary strategies as trait combinations which can stably co-exist \citep{bolker_99}, referring instead to the direction and rate of ongoing adaptive trait evolution under different regimes of stress and disturbance, which is closer in spirit to Grime's arguments \citep{grime_1974,grime_1977}. In addition, we have not assumed any kind of trade-offs or pleiotropy between $b$, $c$ and $d$, only constraints imposed by the environment on the order of magnitude of $b$ and $d$. As an example of a trade-off, corals which rapidly out-shade neighbors have a tall, branched morphology which is vulnerable to disturbances, and so, all else being equal, ideal environment $c$-strategists will suffer higher mortality from disturbances \citep{darling_2012}. Fig.~\ref{fig:axes} gives the same conclusion without invoking trade-offs; mutations which reduce disturbance vulnerability are essentially neutral under ideal conditions, leading to no improvements in mortality from disturbances, whereas $c$ will tend to increase over time. Thus, while trade-offs may amplify specialization, and are sometimes invoked to explain primary strategy schemes \citep{macarthur_1967,winemiller_1992,aerts_1999}, they are not necessary for it.

One potential limitation of our model as a general-purpose model of density-dependent selection is its restriction to interference competition between juveniles for durable resources (lottery recruitment to adulthood), analogous to the ubiquitous assumption of viability selection in population genetics \citep[p. 45]{ewens_2004}. In some respects this is the complement of consumable resource competition models, which restrict their attention to indirect exploitation competition, typically without age structure \citep{tilman_1982}. In the particular case that consumable resources are spatially localized (e.g. due to restricted movement through soils), resource competition and territorial acquisition effectively coincide, and in principle resource competition could be represented by a competitive ability $c$ (or conversely, $c$ should be derivable from resource competition). The situation is more complicated if the resources are well-mixed, since, in general, resource levels then need to be explicitly tracked. It seems plausible that explicit resource tracking may not be necessary when the focus is on the evolution of similar genotypes that use identical resources rather than the stable co-existence of widely differing species with different resource preferences \citep{ram_2016}. We are not aware of any attempts to delineate conditions under which explicit resource tracking is unnecessary even if it is assumed that community structure is ultimately determined by competition for consumable resources. More work is needed connecting resource competition models to the density-dependent selection literature, since most of the former has to date been focused on narrower issues of the role of competition at low resource availability \citep{aerts_1999,davis_1998,tilman_2007}.

@@

In the introduction we mentioned the recurring difficulties with confounding selection and demography in population genetic inference. It seems that Eq. \eqref{eq:master} or something similar (and hopefully more analytically tractable) is unavoidable for the analysis of time-course genetic data because, fundamentally, selective births and deaths affect both abundances and frequencies, not one or the other in isolation. Moreover, some aspects of allele frequency change are intrinsically density-dependent. In the classic lottery model, which as we have seen is essentially the Wright-Fisher model with overlapping generations, $b_i$ and $c_i$ are  equivalent in the sense that the number of territorial victories only depends on the product $b_i c_i$ (see ``Model''). This is no longer the case in our extension, where $b$ and $c$ specialists can co-exist. This ``colonization-competition trade-off'' is well known in the co-existence literature \citep{tilman_94}. It and similar forms of ``spatial co-existence'' in stable environments have previously been modeled either with Levin's  qualitative representation of competition \citep{levins_71,tilman_94}, as opposed to the quantitative $c$ of lottery competition, or with a more sophisticated treatment of space (non-uniform dispersal; \citealt{shmida_84,bolker_99}). In cyclical environments, polymorphisms can be stabilized by the bounded density effect which is  completely lost if there is an exclusive focus on allele frequencies \citep{yi_2013}. We leave the details of how our model might be applied in inference, including the crucial issue of its genetic drift predictions (providing a null model for neutral sites), for future work.

@@

\section*{Appendix C: Mutant invasion under different environments}

In this Appendix we evaluate the invasion of novel mutants in a population with a single resident type $i$ (such that $N=n_i$), which is in equilibrium.

For a single type $i$ in equilibrium ($\Delta n_i = 0$), we have $R_i=0$, $\overline{c}=c_i$, $A_i=(1-(1+L)e^{-L})/L$, and Eq. \eqref{eq:master} becomes
\begin{equation}
b_i(1-e^{-L})/L-d_i=0.\label{eq:equil}
\end{equation}
(Alternatively, Eq. \eqref{eq:equil} can be deduced directly from Eq. \eqref{eq:growthsumuncoupled}). This implies $L\approx b_i/d_i$ when $b_i/d_i\gg 1$ and $L\ll 1$ when $b_i/d_i\approx 1$. Now suppose that a novel mutant $j$, which is initially rare, appears in the population. Then $A_j/R_j\ll 0$, $l_j\approx 0$ and $\overline{c}\approx c_i$, and so, from Eq. \eqref{eq:master}, the mutant lineage's fitness is
\begin{equation}
\Delta n_j/n_j \approx b_j \left(e^{-L}+R_j\frac{c_j}{c_i}\right)-d_j \label{eq:invad}
\end{equation}
where $R_j\approx (1-e^{-L})/\left(\frac{c_j}{c_i}+\frac{L-1+e^{-L}}{1-(1+L)e^{-L}}\right)$ since $l_j\ll 1$.

We now consider mutant invasion under the different environments from section ``Primary strategies and Grime's triangle''. In our representation of Grime's stressful environment, $b\ll 1$, $d\approx b$ and $L\ll 1$, and so Eq. \eqref{eq:invad} becomes $\Delta n_j/n_j \approx b_j-d_j$. Mutations which improve $b$ or $d$ by the same fraction $\epsilon$, such that $b_j = (1+ \epsilon)b_i$ or $d_j = (1- \epsilon)d_i$,  yield identical fitness benefits.

In our representation of Grime's ideal environment, $d\ll 1$, $b_i/d_i \gg 1$ and $L\gg 1$, and so Eq. \eqref{eq:invad} becomes 
\begin{equation}
\Delta n_j/n_j \approx \frac{b_j}{c_j/c_i + L-1}\frac{c_j}{c_i}-d_j\approx \frac{b_j}{L}\frac{c_j}{c_i}-d_j, \label{eq:highLinvad}
\end{equation}
where the last approximation follows from the fact that $c_j$ will be of similar magnitude to $c_i$ (ignoring mutations with very large effect sizes; incidentally, Eq. \eqref{eq:lottery} yields the same expression by assuming $L\rightarrow \infty$ even though this assumption is unrealistic for a rare mutant). Similarly to Grime's stressful environment, improving $b$, $c$, or $d$ by a fraction $\epsilon$ yields identical fitness benefits, even though they may have very different magnitudes, because the ancestral resident type's density-regulated birth rate $b_i c_i/L$ exactly balances its mortality $d_i$ in equilibrium. Biases in the direction of evolution of $b$, $c$ or $d$ must again follow from biases in mutation availability or trade-offs/pleiotropy. The same argument applies for high-$L$ interpretation of the stress regime.

@@

\section*{Appendix D: Coexistence in a stable environment}

To determine whether coexistence is possible in a constant environment, we check for ``mutual invasion'', that is, we check that $j$ will invade an $i$-dominated population, but $i$ will also invade a $j$-dominated population.

We consider the case of coexistence between a $b$ specialist $i$ and a $c$ specialist $j$ ($b_i>b_j$, $c_j>c_i$ and $d_i=d_j$). Suppose that $b_i$ is so large that $L\gg 1$ when $i$ is dominant, and $b_j$ is so small that $L\ll 1$ when $j$ is dominant. Then, when $j$ is dominant, we have $\Delta n_i/n_i=b_i-d_i=b_i-d_j=b_i-b_j>0$. When $i$ is dominant, Eq.  \eqref{eq:highLinvad} applies, where Eq. \eqref{eq:equil} implies $d_j=d_i=b_i(1-e^{-L})/L\approx b_i/L$, and so
\begin{equation}
\Delta n_j/n_j \approx \frac{b_j}{L}\frac{c_j}{c_i}-\frac{b_i}{L}.
\end{equation}
Therefore, coexistence occurs if $c_j/c_i$ is sufficiently large. The analogous argument for $d$ and $c$ specialists ($d_i<d_j$ with $L\gg 1$ when $i$ dominates, $L\ll 1$ when $j$ dominates, and $b_i=b_j$) gives $\Delta n_j/n_j \approx d_i\frac{c_j}{c_i}-d_j$, which again implies coexistence if $c_j/c_i$ is sufficiently large.

For $b$ and $d$ specialists ($c_i=c_j$), we have $\Delta n_j/n_j\approx b_j d_i/b_i-d_j$ when $i$ dominates and $\Delta n_i/n_i\approx b_i d_j/b_j-d_i$ when $j$ dominates. Thus, either $i$ or $j$ grows when rare, but not both, and stable coexistence is not possible in a constant environment.
